---
title: "incarceration"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

### Introduction

url \<-
("<https://raw.githubusercontent.com/melaniewalsh/Neat-Datasets/main/us-prison-pop.csv>")

rl \<- read.csv(url)

View(rl)

Racism still remains as a relevant issue up till today. One area where
its effect could be seen is through the US prison system. incarceration
rates and treatment in criminal justice has shown a systemic
inequalities that dis proportionally effect marginalized people,
particularly people of color. Historical trends have shown that people
of color often received higher rates of imprisonment largely due to
their ethnicity and race, also known as "over policing" on individuals
of color and the war on drug.

The variables I chose for this project is the populations of prisoners
from different ethnicity. Such as white_prison_pop,
black_prison_pop.This will help me gain a better understanding on the
effects of incarceration in different races. In addition I will also be
analyzing the different county city rates, for example their averages
number of imprisonment. This will help me understand whether individuals
from urban or rural areas are disproportionately represented in the
prison population can shed light on how geographic factors intersect
with race and ethnicity within the criminal justice system. other than
that I will be using file 1 to do this.

### Summary Information

Write a summary paragraph of findings that includes the 5 values
calculated from your summary information R script

1.  Proportion_White_Total: 0.379818311105928

```         
white_jail_population <- sum(df$white_jail_pop, na.rm = TRUE)
proportion_white <- white_jail_population / total_jail_population
print(paste("Proportion of Imprisonment for White individuals:", proportion_white))
```

2.  Proportion_Black_Total:0.340023702023979"

```         
total_jail_population <- sum(df$total_jail_pop, na.rm = TRUE)
black_jail_population <- sum(df$black_jail_pop, na.rm = TRUE)
proportion_black <- black_jail_population / total_jail_population
print(paste("Proportion of Imprisonment for Black individuals:", proportion_black))
```

3.  Highest_County: Los Angeles County, 23467.19 Total Pop.

```         
max_index <- which.max(df$total_jail_pop)
Highest_County <- df[max_index, c("county_name", "total_jail_pop")]
```

4.  Lowest_County \<- Clarke County, 0 total pop

```         
min_index <- which.min(df$total_jail_pop)
Lowest_County <- df[min_index, c("county_name", "total_jail_pop")]
print(Lowest_County)
```

5.  Total_Black_Jail\<- 7891239

```         
Total_Black_Jail <- sum(df$black_jail_pop, na.rm = TRUE)
print(Total_Black_Jail)
```

6.  Total_White_Jail\<- 8814789

```         
Total_White_Jail <- sum(df$white_jail_pop, na.rm = TRUE)
print(Total_White_Jail)
```

Summary Findings: Based on the findings above the proportion of
imprisonment for white individuals is approximately 0.3798, meaning that
approximately 37.98% of the total jail population consists of white
individuals. This proportion is interesting in analyzing the extent of
incarceration in US, especially considering the significantly higher
overall white population compared to the black population. However,
despite the stark difference in population of white and black people in
US, the percentage of white individuals in jail remains close to that of
black individuals.

Similarly, the proportion of imprisonment for black individuals is
approximately 0.3400, meaning that approximately 34.00% of the total
jail population consists of black individuals.

The analysis of the highest and lowest county jail populations also
reveals significant disparities. For instance, LA County exhibits a much
higher prison rate than Clarke County with the lowest prison rate. While
this difference could is also due to the vast difference in their
population sizes, it's essential to consider demographic factors.
Despite LA County having a higher Hispanic, and Black population, Clarke
County, with a higher percentage of white residents, has a lower
imprisonment rates overall.

Yet the most interesting findings in this is how the total jail
populations for white and black people remains remarkably similar in
size range. After adding up all the jail populations for white and black
individuals across all states it is evident that white population has
only 1,000,000 higher jail population. According to my research, the
population of white individuals in the US is approximately 231,550,000,
while the population of black individuals is approximately 46,900,000.
Despite white population almost 5 times larger than black population the
prison population is only 1000000 larger. This similarity in prison
rates despite demographic differences highlights the skewed nature of
incarceration in the United States. Showing that races still contributes
to imprisonment rate.

### The Data

Who collected the data?\
Vera Institute. The project was funded by the Robert W. Wilson
Charitable Trust. How was the data collected or generated? The data is
collected by analyzing two Bureau of Justice Statistics (BJS) data
collections, which are the Census of Jails (COJ),and the Annual Survey
of Jails (ASJ). COJ covers all jails and is completed every five to
eight years ever since 1970,and ASJ covers around one-third of jails-and
includes almost all of the largest jails-that has been conducted in
non-census years since 1982.

Why was the data collected?\
Vera use this data in order to depict the jail and prison incarceration
at county level. Before this, incarceration has traditionally centered
on state level data which is effective for understanding the broad
overview of racism in the United States,but lacks the details for
officials-judges, prosecutors, who manage county level imprisonment
causes and consequences. Therefore the Vera data provides with more
detailed over view on incarceration in a county scale, which helps
county level officers gain a better idea of the incarceration and come
up with better consequences and policies.

How many observations (rows) are in your data?\
there are 153,811 rows.

How many features (columns) are in the data?\
There are 36 columns

What, if any, ethical questions or questions of power do you need to
consider when working with this data?\
What are possible limitations or problems with this data? (at least 200
words) When examining these types of data that categorize prison rates
by race and county, ethical and power considerations become relevant
topics. Data regarding race and imprisonment can be very sensitive,
associated with racial stereotypes and biases. When used
unintentionally, these analyses instead of highlighting racism issues
could instead exacerbate discrimination against certain stereotype of
marginalized groups. Therefore, when analyzing this data, it's important
to be sensitive, ensuring that the interpretation doesn't highlight
certain stereotypes or contribute to exacerbating existing racial
marginalization. Moreover, the power dynamics within the criminal
justice system must be acknowledged. These data sets involve systemic
issues like over-policing marginalized communities and disparities in
legal resources, reinforcing societal inequality. Understanding these
contexts will help understand the power imbalances. In addition, there
are also limitations with the data. The data is a broad overview of the
number of races but lacks data and detail about the circumstances of
each case. Since we mostly only deal with numerical data and don't have
much access to the detailed overview of case by case, the conclusions
may not be fully accurate. Interpreting raw incarceration rates can lead
to wrong conclusions about the fairness and criminal justice system.
Maybe in order for more accurate analysis, more detailed case overview
could give a more fair idea. With more information regarding prison
reasoning, socioeconomic factors, access to education, and healthcare
may lead to better conclusions regarding incarceration rates. Therefore,
the analysis should be approached with careful consideration and must be
taken into consideration of other considerations as well from outsource.

## Graph 1: Trends Over Time of Prison Population by Race

I've included this stack bar graph to compare prison populations across
different racial groups. Based on the graph it's evident that the Latino
population has the highest representation in prisons. This observation
reinforces the relationship between race and prison population rates.
The bar graph effectively highlights this disparity over the years,
providing a clear visual comparison. Furthermore, with the X axis
showing the years, the graph also shows the progression on incarceration
over the years.

```{r CHART1_PALMY, echo = FALSE}
source("CHART1_PALMY.R")
plot(g1)
```

\`\`\``{r, echo = FALSE, code = readLines("chart1_example.R")}`

## Graph 2: Relationship Between Total Population and Prison Population

I have included a scatter plot to analyze the relationship between total
population and prison population across various states. The graph
depicts a positive correlation, indicating that as the population
increases, so does the prison population, which is expected. However,
there are a few outlier that either surpass or fall below this
trend.When examining this relationship, we can gain insights into the
extent of incarceration within a population and how it scales with
population growth.

```{r CHART2, echo = FALSE}
source("CHARTNUMBER2.R")
print(p2)
```

## Map in America Patterns Population

I use map to model the total prison population across different states
through the different color gradient. By modelling the population of
states we can considering not only population numbers but also
demographic composition. This approach can help uncover correlations
between racial demographics and incarceration rates. For instance,
states with higher minority populations might have higher rates of
imprisonment, which may reflect a potential systemic biases or racial
disparities within the criminal justice system

```{r CHARTNUMBER3, echo = FALSE}
source("CHARTNUMBER3copy")
print(p5)

```

